Linear Stochastic Approximation- Constant Step-Size and Iterate Averaging

### 2017-05-11 intro.tex
revisions

\section{Introduction}
A linear stochastic approximation (LSA) algorithm is given by
\begin{align}\label{eq:lsaintro}
\theta_t=\theta_{t-1}+\alpha_t (b_t-A_t \theta_{t-1}),
\end{align}
where the iterates, $\theta_t\in \R^d$, and step-sizes, $\alpha_t>0$, are algorithmic quantities, and $(b_t,A_t)\in \R^d\times \R^{\dcd}$ is the noisy data sequence presented to the algorithm.
LSA algorithms are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using the noisy data sequence, using $O(d)$ updates per iteration. Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear prediction with quadratic loss \cite{bach,bachaistats}, and the temporal difference class (TD) of learning algorithms in RL \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.\par
%An additional feature in these class of applications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par
It is of interest to understand the effect of the step-sizes $\alpha_t$ (a quantity that can be controlled by the algorithm designer) on the performance of the LSA in \eqref{eq:lsaintro}. The choice of the step-sizes dictates learning rate, noise rejection and stability. Informally speaking, smaller step-sizes are better for noise rejection and larger step-sizes lead to faster learning, while at the same time too large step-sizes might result in instability of \eqref{eq:lsaintro}. Diminishing and square-summable step-sizes have been a popular choice for the step-sizes, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$ \cite{gtd2,borkarbook,konda-tsitsiklis}. An alternative technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{s=0}^t \theta_t$. In LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm.
Averaging in stochastic approximation to improve noise rejection is a procedure that goes back to  Polyak and Ruppert \cite{polyak-judisky,ruppert}.
%In what follows, when we say CS-PR for LSA or LSA with CS-PR, we allude to the case where \eqref{eq:lsaintro} is a constant step-size LSA with $\alpha_t=\alpha>0,\forall t\geq 0$, $\theta_t$ in \eqref{eq:lsaintro} is considered an internal variable and $\thh_t$ is the final output of the LSA with CS-PR.
%In LSA with CS-PR, \eqref{eq:lsaintro} is a constant step-size LSA with $\alpha_t=\alpha>0,\forall t\geq 0$, $\theta_t$ in \eqref{eq:lsaintro} is considered an internal variable and $\thh_t$ is the final output of the LSA with CS-PR.
Intuitively, the CS-PR technique can be seen to address the learning vs noise rejection trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the noise rejection. \par
Recently, \cite{bach} considered SGD\footnote{SGD is a special of the LSA algorithm in \eqref{eq:lsaintro}.} with CS-PR for the linear prediction problem with quadratic loss and $i.i.d$ sampling. A ``remarkable'' result by \cite{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the expected squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\par\todoc{Dimension?}
\paragraph{Motivation and Contribution:} In this paper, we wish to study LSA with CS-PR for $i.i.d$ data with bounded variance. An additional restriction we place is that the matrix $\EEP{A_t}$ (where $A_t\sim P$ ) is Hurwitz, i.e., all its eigenvalues have positive real parts. For a given $P$, we wish to measure the performance in terms of the mean square error (MSE) $\EEP{\norm{\thh_t-\ts}^2}$. Further, for a given class of distributions $\P$, we also would like to know whether it is possible to ensure uniform performance $\forall P\in \P$.
Our study is motivated by the problem \emph{policy evaluation} \cite{dann} using linear value function approximation from experience replay \cite{lin} in a batch setting \cite{lange} in RL using TD class \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp} of algorithms. A result on the performance of LSA with CS-PR can be directly translated for the TD class of algorithms that are also LSA algorithms.
We now list our contributions in this paper
%While asymptotically diminishing step-sizes are common in RL literature \cite{ilstd,gtdmp,korda-prashanth}, the constant step-size choice has also been used in the past \cite{gtd2}.
\begin{itemize}[leftmargin=*]%, before = \leavevmode\vspace{-\baselineskip}]
\item \textbf{Instance Dependent Bounds} We show that for a fixed $P$, there exists a $\alpha_P>0$ such that the mean squared error (MSE) of the LSA with CS-PR is at most $O(\frac{1}{t})$, whenever $\alpha\in (0,\alpha_P)$. The MSE further can be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
\item \textbf{Uniform Step-Sizes:} We also show (via counter example) that there is a class $\P$ that does not `admit' a constant step-size choice that guarantees uniform performance $\forall P\in \P$. This is a negative result in the context of RL. However,
we also argue that the subclass $\P_{LR}$ related to SGD for linear regression does `admit' a problem independent constant step size, thereby making a connection between our framework and the one considered by \cite{bach}.
\item \textbf{Automatic Step-Size:} We argue informally that for any given $P$, a constant step-size can be chosen in an `automated' fashion, and the performance of CS-PR is comparable to other popular step-size choices used for LSA in RL literature.
\end{itemize}
\paragraph{Organization:}

>>>>

\section{Introduction}
A linear stochastic approximation (LSA) algorithm is given by
\begin{align}\label{eq:lsaintro}
\theta_t=\theta_{t-1}+\alpha_t (b_t-A_t \theta_{t-1}),
\end{align}
where the iterates, $\theta_t\in \R^d$, and step-sizes, $\alpha_t>0$, are algorithmic quantities, and $(b_t,A_t)\in \R^d\times \R^{\dcd}$ is the noisy data sequence presented to the algorithm.
LSA algorithms (or simply LSAs) are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using only the noisy data sequence and $O(d)$ updates per iteration. Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear regression with quadratic loss \cite{bach,bachaistats}, and the temporal difference class (TD) of learning algorithms in RL \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.\par
%An additional feature in these class of applications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par

The choice of the step-sizes $\alpha_t$ is critical for the performance of the LSA in \eqref{eq:lsaintro}.
Informally speaking, smaller step-sizes are better for noise rejection and larger step-sizes lead to faster learning, while at the same time too large step-sizes might result in instability of \eqref{eq:lsaintro}. A useful choice in RL has been the diminishing and square-summable step-sizes \cite{gtd2,borkarbook,konda-tsitsiklis}, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$. An alternate technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{s=0}^t \theta_t$. In LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm. The idea of averaging in stochastic approximation to improve noise rejection is a technique that goes back to  \citet{ruppert,polyak-judisky}. Intuitively, the CS-PR technique can be seen to address the learning vs noise rejection trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the noise rejection.
\par
Recently, \citet{bach} considered SGD\footnote{SGD is a special of the LSA algorithm in \eqref{eq:lsaintro}.} with CS-PR for the linear regression problem with quadratic loss and $i.i.d$ sampling. A ``remarkable'' result by \citet{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the expected squared regression error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\par\todoc{Dimension?}

\paragraph{Motivation and Contribution:} In this paper, we wish to study LSA with CS-PR where data $(b_t,A_t)$ is $i.i.d$ with some underlying distribution $P$ such that variance is bounded. An additional restriction we place on $P$ is that the matrix $\EEP{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts. For a given $P$, we wish to measure the performance of LSA with CS-PR in terms of the mean square error (MSE) $\EEP{\norm{\thh_t-\ts}^2}$. Further, for a given class of distributions $\P$, we would like to know whether it is possible to ensure uniform performance $\forall P\in \P$.
Our study is motivated by the problem \emph{policy evaluation} \cite{dann} using linear value function approximation from experience replay \cite{lin} in a batch setting \cite{lange} in RL using TD class  of algorithms \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}. Here, we wish to understand the effectiveness of the CS-PR technique for the TD class of LSA algorithms used in RL.

We now list our contributions in this paper
%While asymptotically diminishing step-sizes are common in RL literature \cite{ilstd,gtdmp,korda-prashanth}, the constant step-size choice has also been used in the past \cite{gtd2}.
%\begin{itemize}[leftmargin=*]%, before = \leavevmode\vspace{-\baselineskip}]
 \begin{center}Algorithm\end{center}
\paragraph{Instance Dependent Bounds} We show that for a fixed $P$, there exists a $\alpha_P>0$ such that the mean squared error (MSE) of the LSA with CS-PR is at most $O(\frac{1}{t})$, whenever $\alpha\in (0,\alpha_P)$. The MSE can further be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
\begin{center} Structure in Data\end{center}
\paragraph{Uniform Step-Sizes:} We also show (via counter example) that there is a class $\P$ that does not `admit' a single constant step-size $\alpha_{\P}$ that guarantees uniform performance $\forall P\in \P$. This is a negative result in the context of RL. However,
we also argue that the subclass $\P_{LR}$ related to SGD for linear regression does `admit' a problem independent constant step-size, thereby recovering a part of the result by \citet{bach}.
\begin{center} Meta-Parameter $\alpha$\end{center}
\paragraph{Automatic Step-Size:} We argue informally that for any given $P$, a constant step-size can be chosen in an `automated' fashion, and the performance of CS-PR is comparable to other popular step-size choices used for LSA in RL literature.
\paragraph{Organization:}

### 2017-05-11 not.tex
revisions

\section{Notations and Definitions}\label{sec:def}

We denote the sets real and complex number by $\R$ and $\C$ respectively. For $x=a+ib\in \C$, we denote its real and imaginary parts by $\re{x}=a$ and $\im{x}=b$ respectively. We denote $n$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$ respectively. We use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries respectively. For any $x\in \C$ we denote the modulus of $x$ by $\md{x}=\sqrt{\re{x}^2+\im{x}^2}$. For any $x\in \C^d$We denote complex conjugate of $x$ by $\bar{x}$. We denote the transpose of $x\in \C^{d}, A\in \C^{\dcd}$ by $x^\top$ and $A^\top$ respectively. Further, by $x^*$ denotes the conjugate transpose, i.e., $x^*=\bar{x}^\top$ (the same convention carries over to matrices as well). We will use $\ip{\cdot,\cdot}$ to denote the inner products: $\ip{x,y}=x^* y$. Also, by $A_{ij}$ we denote the entry of matrix $A$ in the $i^{th}$ row and $j^{th}$ column.
The norm of the matrix $A$ in given by $\norm{A}\eqdef \sup_{x\in \C^d:\norm{x}=1} \norm{Ax}$. For a positive real number $B>0$, we denote by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \norm{A}\leq B\}$ and  $\R^{\dcd}_B=\{A\in \R^{\dcd}\mid \norm{A}\leq B,\,\forall i,j=1,\ldots,n\}$ the set of complex and real matrices whose norms are bounded $B$.
%We denote by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$ and  $\R^{\dcd}_B=\{A\in \R^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$.
%We use $\P$ to denote the set of probability distribution over $\C^{\dcd}$.
We also denote the set of invertible $\dcd$ complex matrices by $\gln$. We use $A\sim P$ to denote the fact that the random variable $A$ is distributed according to distribution $P$. Let the random matrix $A\sim P$, and $U\in \gln$, we use $P_U$ to denote the distribution of the random matrix $U^{-1}A U$.  We denote the identity matrix in $\C^{\dcd}$ by $\I$.
%We denote the set of distributions over $\C_B^{\dcd}$ by $\P$.

We use $A\succeq 0$ to denote that the
square matrix $A$ is hermitian and positive semidefinite (HPD):
$A = A^*$, $\inf_x x^* A x\ge 0$.
For $A,B$ HPD matrices, $A\succeq B$ holds if $A-B\succeq 0$.
We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
We also use $\preceq$ and $\prec$ analogously. For any $x\in\C^d$x, we denote the general quadratic norm with respect to a $C\succ 0$ by $\norm{x}^2_C\eqdef x^*\, C \,x$.
$\E$ denotes mathematical expectation.

\begin{definition}
Let $P$ be a distribution over $\C^{\dcd}$, then define
\begin{align*}
A_P=\int M dP(M),\quad C_P=\int M^* M dP(M),
\end{align*}
\end{definition}
\begin{definition}\label{def:contract}[Contraction Factors]
Define
\begin{align*}
\rhod{P}&\eqdef {\inf}_{x\in\C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha A_P^* A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha C_P\right)x}
\end{align*}
\end{definition}
Note that $\rhod{P}>\rhos{P}$.

\begin{definition}
We call a matrix $A\in \C^{\dcd}$ to be \emph{positive definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{d}$. We call a matix $A\in \C^{\dcd}$ to be \emph{asymptotically stable} (AS) if all eigenvalues of $A$ have positive real parts. We call a matrix $A\in \C^{\dcd}$ to be \emph{symmetric positive definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD.
\end{definition}
\begin{definition}\label{distpd}
We call a distribution $P$ to be PD/AS/SPD if $A_P$ is PD/AS/SPD.
\end{definition}
\begin{example}
The matrices $\begin{bmatrix}0.1 &-1\\ 1 & 0.1\end{bmatrix}$, $\begin{bmatrix} 0.1 & 0.1 \\ 0 & 0.1\end{bmatrix}$ and $\begin{bmatrix}0.1 &0 \\ 0 & 0.1\end{bmatrix}$ are examples of AS, PD and SPD respectively.
\end{example}

>>>>

\section{Notations and Definitions}\label{sec:def}
We denote the sets real and complex numbers by $\R$ and $\C$ respectively. For $x=a+ib\in \C$, we denote its real and imaginary parts by $\re{x}=a$ and $\im{x}=b$ respectively.
For any $x\in \C$ we denote its modulus $\md{x}=\sqrt{\re{x}^2+\im{x}^2}$ and its complex conjugate by $\bar{x}=a-ib$.
We denote $d$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$ respectively.
We denote the transpose of $x\in \C^{d}$ by $x^\top$ and the conjugate transpose by $x^*={\bar{x}}^\top$.
We will use $\ip{\cdot,\cdot}$ to denote the inner products: $\ip{x,y}=x^* y$.
%We denote the transpose of $x\in \C^{d}, A\in \C^{\dcd}$ by $x^\top$ and $A^\top$ respectively.
%Further, by $x^*$ denotes the conjugate transpose, i.e., $x^*=\bar{x}^\top$ (the same convention carries over to matrices as well).

 %Also, by $A_{ij}$ we denote the entry of matrix $A$ in the $i^{th}$ row and $j^{th}$ column.
We use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries respectively. The notation for (conjugage) transpose carries over from vectors.
The norm of the matrix $A$ in given by $\norm{A}\eqdef \sup_{x\in \C^d:\norm{x}=1} \norm{Ax}$. For any $x\in\C^d$, we denote the general quadratic norm with respect to a hermitian matrix $C$ by $\norm{x}^2_C\eqdef x^*\, C \,x$. We use $\cond{A}=\norm{A}\norm{A^{-1}}$ to denote the condition number of the matrix $A$. We denote the identity matrix in $\C^{\dcd}$ by $\I$ and the set of invertible $\dcd$ complex matrices by $\gln$.
\begin{comment}
We use $A\succeq 0$ to denote that the
square matrix $A$ is hermitian and positive semidefinite (HPD):
$A = A^*$, $\inf_x x^* A x\ge 0$.
For $A,B$ HPD matrices, $A\succeq B$ holds if $A-B\succeq 0$.
We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
We also use $\preceq$ and $\prec$ analogously.
\end{comment}
For a positive real number $B>0$, we denote by $\C^{d}_B=\{b\in \C^d\mid \norm{b}\leq B\}$ and by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \norm{A}\leq B\}$ the set of complex vectors and matrices whose norms are bounded $B$.
%We denote by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$ and  $\R^{\dcd}_B=\{A\in \R^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$.
%We use $\P$ to denote the set of probability distribution over $\C^{\dcd}$.
%We denote the set of distributions over $\C_B^{\dcd}$ by $\P$.
Let $P$ be a distribution over $C^{\dcd}$, we use $A\sim P$ to denote $A$ is distributed according to $P$. For $U\in \gln$, we denote by $P_U$ the distribution of $U^{-1}\,A\,U$, where $A\sim P$. $\E$ denotes mathematical expectation.
\begin{comment}
 We use $A\sim P$ to denote the fact that the random variable $A$ is distributed according to distribution $P$. Let the random matrix $A\sim P$, and $U\in \gln$, we use $P_U$ to denote the distribution of the random matrix $U^{-1}A U$.
\end{comment}
\begin{definition}\label{def:dist}
Let $P=(P^V,P^M)$ denote a $2$-tuple of probability distributions; $P^V$ over $\C^{d}$ and $P^M$ over $\C^{\dcd}$. Define
\begin{align*}
A_P&=\int M dP^M(M),\quad C_P=\int M^* M dP^M(M), \quad b_P=\int V dP^V(V)\\
\rhod{P}&\eqdef {\inf}_{x\in\C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha A_P^* A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha C_P\right)x}
\end{align*}
\end{definition}
Note that $\rhod{P}>\rhos{P}$.
\begin{definition}\label{def:simdist}
Let $P=(P^V,P^M)$ as in \Cref{def:dist}; $b\sim P^V$ and $A\sim \P^M$ be random variables distributed according to $P^V$ and $P^M$. For $U\in \gln$ define $P_U=(P_U^V,P_U^M)$ to be the distributions of $U^{-1}b$ and $U^{-1}AU$.
\end{definition}
\begin{definition}
We call a matix $A\in \C^{\dcd}$ to be \emph{Hurwitz} (H) if all eigenvalues of $A$ have positive real parts. We call a matrix $A\in \C^{\dcd}$ to be \emph{Positive Definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{d}$.  We call a matrix $A\in \C^{\dcd}$ to be \emph{Symmetric Positive Definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD.
\end{definition}
\begin{definition}\label{distpd}
We call a distribution $P$ to be H/PD/SPD if $A_P$ is H/PD/SPD.
\end{definition}
\begin{example}
The matrices $\begin{bmatrix}0.1 &-1\\ 1 & 0.1\end{bmatrix}$, $\begin{bmatrix} 0.1 & 0.1 \\ 0 & 0.1\end{bmatrix}$ and $\begin{bmatrix}0.1 &0 \\ 0 & 0.1\end{bmatrix}$ are examples of H, PD and SPD respectively.
\end{example}

\begin{definition}
Call a set of distributions $\P=(\P^V,\P^M)$ over $\C^{d}\times \C^{\dcd}$
\emph{weakly admissible} if there exists $\alpha_{\P}>0$ such that
$\rhos{P}>0,\forall \alpha\in(0,\alpha_{\P})$ holds for all $P\in \P$.
\end{definition}
\begin{definition}
Call a set of distributions $\P=(\P^V,\P^M)$ over $\C^{d}\times \C^{\dcd}$ \emph{admissible}
if there exists some $\alpha_{\P}>0$ such that $\inf_{P\in \P} \rhos{P}>0,\,\forall \alpha\in(0,\alpha_{\P})$.
The value of $\alpha_{\P}$ is called a witness.
\end{definition}

It is easy to see that $\alpha \mapsto \rhos{P}$ is decreasing,
hence if $\alpha_{\P}>0$ witnesses that $\P$ is (weakly) admissible
then any $0<\alpha'\le \alpha_{\P}$ is also witnessing this.

### 2017-05-11 prob.tex
revisions

We consider the linear stochastic approximation (LSA) algorithm with constant step-size (CS) and Polyak-Ruppert (PR) averaging of the iterates.

@@

can be implemented in $O(d)$ time and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.\par
We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa}. In what follows, we make use of the $\sigma$-field given by $\F_{t-1}=\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.

@@

\item \label{dist} $\{A_t\in\C^{\dcd},\, t\geq 0\}$ is an $i.i.d$ sequence distributed according to $P$ which is either AS or PD or SPD.
\item $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^d$ with mean $\E[b_t]=b$.
q\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ associated with $A_t$ and $b_t$ satisfy the following
%\begin{align*}\E\left[ M_t^* M_t\mid\F_{t-1}\right]\leq \Sigma^2_P, \, \E[N_t^* N_t]=\sigma_b^2.\end{align*}
\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_P, \, \E[N_t^* N_t]=\sigma_b^2.\end{align*}
\item $A_P$ is invertible and there exists a $\ts=A_P^{-1}b$.

@@

\paragraph{Performance Metric:}  We are interested in the behavior of the mean squared error (MSE) at time $t$ given\footnote{ We note here that $\EE{\norm{\thh_t-\ts}_C^2}$ which measures the MSE in the quadratic norm with respect to a positive definite matrix $C\succ 0$. However, unless $C$ has some special structure, it is not possible to exploit this generality in our bounds. Further, when data is bounded $\norm{C}^2$ is also bounded and it is straightforward to see that $\EE{\norm{\thh_t-\ts}}^2_C\leq \norm{C}^2\EE{\norm{\thh_t-\ts}^2}$. So, for the purpose of stating our results we will not use the general quadratic norm, however, will use it when we discuss the work by \cite{bach} in the linear regression setting, where its usage is relevant.} by $\EE{\norm{\thh_t-\ts}^2}$. We wish to understand how the MSE is affected by the choice of $\alpha$ and the spectral properties of $A_P$. Another aspect of interest is to understand the contribution of the \emph{bias} (caused by the initial condition $\theta_0$) and the \emph{variance} (caused by $\sigma^2_P,\sigma^2_b$) to the MSE.\par

>>>>

We consider the linear stochastic approximation (LSA) algorithm with constant step-size (CS) and Polyak-Ruppert (PR) averaging of the iterates given as below:

@@

can then be implemented in $O(d)$ time and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.\par
We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa}. In what follows, we make use of the $\sigma$-field given by $\F_{t-1}\eqdef\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.

@@

\item \label{dist} $(b_t, A_t)\sim (P^b,P^A)$ is an $i.i.d$ sequence, where $P^b$ is a distribution over $\C^d$ and $P^A$ is a Hurwitz distribution over $\C^{\dcd}$.
\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following
\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_{A_P}, \, \E[N_t^* N_t]=\sigma^2_{b_P}.\end{align*}
\item $A_P$ is invertible and there exists a $\ts=A^{-1}_Pb_P$.

@@

\paragraph{Performance Metric:}  We are interested in the behavior of the mean squared error (MSE) at time $t$ given\footnote{ $\EE{\norm{\thh_t-\ts}_C^2}$, the MSE with respect a hermitian matrix $C$ is desirable, we cannot exploit the use of $C$ unless it is connected to $P$ in a special way. For more discussion, see \Cref{sec:related}.}
 by $\EE{\norm{\thh_t-\ts}^2}$.
\footnote{ We note here that $\EE{\norm{\thh_t-\ts}_C^2}$ which measures the MSE in the quadratic norm with respect to a positive definite matrix $C\succ 0$. However, unless $C$ has some special structure, it is not possible to exploit this generality in our bounds. Further, when data is bounded $\norm{C}^2$ is also bounded and it is straightforward to see that $\EE{\norm{\thh_t-\ts}}^2_C\leq \norm{C}^2\EE{\norm{\thh_t-\ts}^2}$. So, for the purpose of stating our results we will not use the general quadratic norm, however, will use it when we discuss the work by \cite{bach} in the linear regression setting, where its usage is relevant.}
%We wish to understand how the MSE is affected by the choice of $\alpha$ and the spectral properties of $A_P$. Another aspect of interest is to understand the contribution of the \emph{bias} (caused by the initial condition $\theta_0$) and the \emph{variance} (caused by $\sigma^2_P,\sigma^2_b$) to the MSE.\par
